\section{Introduction} \label{introduction} 

Progress in some branches of
mathematics might be restricted by size of symbolic derivations that human can
handle. Mathematicians can deal with expressions having dozens of
variables but would struggle to discover useful relations between hundreds or
thousands of variables. We will show how the task of finding equivalent
mathematical expressions can be automated. Our focus is on finding equivalent
mathematical formulas (i.e.~which give an identical numerical result
to the original expression),
but are faster to compute. The definition
of faster can be either (i) the number of operations or (ii) the computational time
for particular hardware.

We propose a deterministic, grammar based framework, which discovers
relations between multi-variable polynomial expressions. First, we
construct an attribute grammar -- a context-free grammar extended to
contain set of attributes, introduced by Donald Knuth
\cite{knuth1968semantics}. To define the search space we use a set of
context-free grammar rules representing admissible operations. By
representing the cost of every operation as a synthesized attribute
(i.e. computed bottom-up from child node attributes) we can search for
formula with low time complexity. Through a linear combination of
grammar elements, we can find a solution to the desired
expressions. Finally, this computation solution can be further speed
up by use of standard techniques of optimization in compiler
domain. We show the power of the approach by deriving a closed form,
$O(n^3)$ time solution (where $n$ is the number of units) for a Taylor
series approximation to the partition function of an RBM. This
violates the conventional wisdom that accurate computation of the
partition function requires an exponential number of elements.


XXX : Change according to Joan's section.
Finally, we can regard the set of generated rules as a small axiomatic system.
The presented algorithm is deterministic, and provably runs in finite
time. We are excited about using machine learning for automatic reasoning, and
this subset of mathematics seems to be perfect fit for testing automatic reasoning systems.
Usually, axiomatic systems like number theory, set theory, or topology have very small
number of axioms, but their application to terms might be complex. Proofs
for such systems are difficult to represent in software, and there is no
baseline algorithms which could brute force proof by iterating over all of them in reasonable finite time.
Where in the contrary, presented here axiomatic system has very simple representation
in software, and this paper presents brute force algorithm to find proofs. 
Surprisingly, even brute force algorithm is able to find new more efficient,
computational expressions for expressions that we use.

%% ROB: Put at end, I think.
%By the time of publication we are going to release code.



